---
title: 'Using Azure Data Science Virtual Machine: Introduction'
author: "Le Zhang and Graham Williams"
date: '`r Sys.Date()`'
output: rmarkdown::html_vignette
vignette: >
 %\VignetteIndexEntry{Vignette Title}
 %\VignetteEngine{knitr::rmarkdown}
 \usepackage[utf8]{inputenc}
---

The Azure Data Science Virtual Machine (DSVM) provide a powerful
platform for data scientists supporting a full open source stack of
tools that today's data scientist will regularly use.
	
# Preliminaries

1. `AzureDSVM` requires users to have access to Azure resources
through an Azure subscription which can be obtained through a [free
Azure account](https://azure.microsoft.com/en-us/free).

2. It is highly recommended that you read through the DSVM
documentation:

      * [An Overview of the DSVM](https://docs.microsoft.com/en-us/azure/machine-learning/machine-learning-data-science-virtual-machine-overview);
      * [Ten things you can do on the DSVM](https://docs.microsoft.com/en-us/azure/machine-learning/machine-learning-data-science-vm-do-ten-things);
      * [DSVM GitHub repository](https://github.com/Azure/Azure-MachineLearning-DataScience/tree/master/Misc/DataScienceProcess/DataScienceScripts/Solution%20Arch/DSVM).

3. `AzureDSVM` assumes R (>= 3.3.1), with dependencies on packages of
`AzureSMR `(>= 0.2.2), `stringr` (>= 1.1.0), `stringi` (>= 1.1.2),
`magrittr` (>= 1.5), `dplyr` (>= 0.5.0). `AzureSMR` can be installed
from its [GitHub repository](https://github.com/Microsoft/AzureSMR) as
described below. The other packages are available on CRAN or MRAN.

4. Before using the functions in `AzureDSVM`, one needs to obtain
authentication credentials for managing Azure resources. Details are
available from the [AzureSMR Authentication
vignette](https://github.com/Microsoft/AzureSMR/blob/master/vignettes/Authentication.Rmd).

# How-to

## Installation

`AzureDSVM` is available on [GitHub repository](https://github.com/Azure/AzureDSVM). To install, 

```{r install azuresvm, eval=FALSE}
if(!require("devtools")) install.packages("devtools")
devtools::install_github("Azure/AzureDSVM")
```

Help pages can be loaded by

```{r help azuresvm, eval=FALSE}
library(help=AzureDSVM)
```

## Tutorials

Tutorials on how to use `AzureDSVM` package are provided in `/vignettes`.

* [Get started](https://github.com/Azure/AzureDSVM/blob/master/vignettes/00Introduction.Rmd)
* [Deployment of a single DSVM](https://github.com/Azure/AzureDSVM/blob/master/vignettes/10Deploy.Rmd)
* [Deployment of multiple DSVMs](https://github.com/Azure/AzureDSVM/blob/master/vignettes/20Multi.Rmd)
* [Do computation on a single DSVM or a cluster of DSVMs](https://github.com/Azure/AzureDSVM/blob/master/vignettes/30Compute.Rmd)
* [Monitor data consumption and expense spent on using DSVM(s)](https://github.com/Azure/AzureDSVM/blob/master/vignettes/40Cost.Rmd)
* [Putting all together - use case of kmeans clustering](https://github.com/Azure/AzureDSVM/blob/master/vignettes/60Kmeans.Rmd)
* [Putting all together - use case of binary classification](https://github.com/Azure/AzureDSVM/blob/master/vignettes/80ModelSelect.Rmd)
