---
title = "Using Azure Data Science Virtual Machine: Deployment of a single DSVM"
author= "Graham Williams"
---

# Use Case

In this tutorial, both a Linux and a Windows Data Science Virtual Machine (DSVM) are deployed. 
Code is included but
not run to then delete the resource group if the resources are no
longer required. Once deleted consumption will cease.

This script is best run interactively to review its operation and to
ensure that the interaction with Azure completes. As a standalone
script it can be run to setup a new resource group and single Linux
DSVM.

# Preparation

We assume the user already has an Azure subscription and we have
obtained the credentials required. See instructions in the [previous vignette](https://github.com/Azure/AzureDSVM/blob/master/vignettes/00Introduction.Rmd) 
for details. We will then ensure a resource group exists and within
that resource group create DSVMs. A public ssh key is used to
access the server in this script although a username and password is
also an option. We create a Linux DSVM and a Windows DSVM.

# Setup

To get started we need to load our Azure credentials as well as the
user's ssh public key. Public keys on Linux are typically created on
the users desktop/laptop machine and will be found within
~/.ssh/id_rsa.pub. It will be convenient to create a credentials file
to contain this information. The content's of the credentials file
will be something like:

```{r credentials, eval=FALSE}
# Credentials come from app creation in Active Directory within Azure.
 
TID <- "72f9....db47"          # Tenant ID
CID <- "9c52....074a"          # Client ID
KEY <- "9Efb....4nwV....ASa8=" # User key

PUBKEY   <- readLines("~/.ssh/id_rsa.pub") # For Linux DSVM
PASSWORD <- "Public%4aR3@kn"               # For Windows DSVM

```

Save such information into a file with the name <USER>_credentials.R
where <USER> is replaced with your username. Then we simply source
that file in R.

```{r setup}
# Load the required subscription resources: TID, CID, and KEY.
# Also includes the ssh PUBKEY for the user.

USER <- Sys.info()[['user']]

source(paste0(USER, "_credentials.R"))
```

```{r, eval=FALSE}
# Install the packages if required.

devtools::install_github("Microsoft/AzureSMR")
devtools::install_github("Azure/AzureDSVM", auth_token=GIT_TOKEN) # auth_token is only required when repo is private.
```

```{r packages}
# Load the required packages.

library(AzureSMR)    # Support for managing Azure resources.
library(AzureDSVM)   # Further support for the Data Scientist.
library(magrittr)    
library(dplyr)
```

```{r tuning}
# Parameters for this script: the name for the new resource group and
# its location across the Azure cloud. The resource name is used to
# name the resource group that we will create transiently for the
# purposes of this script.

# Create a random resource group to reduce likelihood of conflict with
# other users.

BASE <- 
  runif(4, 1, 26) %>%
  round() %>%
  letters[.] %>%
  paste(collapse="") %T>%
  {sprintf("Base name:\t\t%s", .) %>% cat("\n")}

RG <-
  paste0("my_dsvm_", BASE,"_rg_sea") %T>%
  {sprintf("Resource group:\t\t%s", .) %>% cat("\n")}

# Choose a data centre location.

LOC <-
  "southeastasia"  %T>%
  {sprintf("Data centre location:\t%s", .) %>% cat("\n")}

# Include the random BASE in the hostname to reducely likelihood of
# conflict.

HOST <-
  paste0("my", BASE) %T>%
  {sprintf("Hostname:\t\t%s", .) %>% cat("\n")}

cat("\n")
```

```{r connect}
# Connect to the Azure subscription and use this as the context for
# our activities.

context <- createAzureContext(tenantID=TID, clientID=CID, authKey=KEY)

# Check if the resource group already exists. Take note this script
# will not remove the resource group if it pre-existed.

rg_pre_exists <- existsRG(context, RG, LOC)
```
# Create a Resource Group

Create the resource group within which all resources we create will be
grouped.

```{r create resource group}
if (! rg_pre_exists)
{
  # Create a new resource group into which we create the VMs and
  # related resources. Resource group name is RG. 
  
  # Note that to create a new resource group one needs to add access
  # control of Active Directory application at subscription level.

  azureCreateResourceGroup(context, RG, LOC) %>% cat("\n\n")

}

# Check that it now exists.

cat("Resource group", RG, "at", LOC,
    ifelse(!existsRG(context, RG, LOC), "does not exist.\n", "exists.\n"), "\n")
```

# Deploy a Linux Data Science Virtual Machine

Create the actual Linux DSVM with public-key based authentication
method. Name, username, and size can also be configured.

Note one can check available VM sizes within the region by using `getVMSizes()`. Different sizes will cost differently, and the detailed information can be checked on [Azure website](https://docs.microsoft.com/en-us/azure/virtual-machines/virtual-machines-windows-sizes). The default VM size for deployment is "Standard_DS2_v2", which is featured by enhanced computation performance.

```{r}
getVMSizes(context, "southeastasia")
```

The following code deploys a Linux DSVM, and it will take approximately 4 minutes.

```{r deploy}
# Create the required Linux DSVM - generally 4 minutes.

ldsvm <- deployDSVM(context, 
                    resource.group=RG,
                    location=LOC,
                    hostname=HOST,
                    username=USER,
                    pubkey=PUBKEY)
ldsvm

operateDSVM(context, RG, HOST, operation="Check")

azureListVM(context, RG)
```

Prove that the deployed DSVM exists.

```{r prove exists}

# Send a simple system() command across to the new server to test its
# existence. Expect a single line with an indication of how long the
# server has been up and running.

cmd <- paste("ssh -q",
             "-o StrictHostKeyChecking=no",
             "-o UserKnownHostsFile=/dev/null",
             ldsvm, "uptime")
cmd
system(cmd, intern=TRUE)

```

# Deploy a Windows Data Science Virtual Machine

`deployDSVM` also supports deployment of Windows DSVM, which can be
achieved by setting the argument of `os` to "Windows". The deployment will take approximately 10 minutes. One can use Remote Desktop to verify the success of deployment and use the virtual machine in a remote desktop environment.

```{r, eval=FALSE}
wdsvm <- deployDSVM(context,
                    resource.group=RG,
                    location=LOC,
                    hostname="xxxx",
                    username=USER,
                    os="Windows",
                    password=PASSWORD)

wdsvm
```

# Optional Stop

It is always a good practice to stop DSVMs after using them, to avoid any unnecessary cost.

```{r, eval=FALSE}
operateDSVM(context, RG, HOST, operation="Stop")
```

# Optional Cleanup

```{r optionally delete resource group}
# Delete the resource group now that we have proved existence. There
# is probably no need to wait. Only delete if it did not pre-exist
# this script. Deletion seems to take 10 minutes or more.

if (! rg_pre_exists)
  azureDeleteResourceGroup(context, RG)
```

Once deleted we are consuming no more.
