---
title  = "Using Azure Data Science Virtual Machine: Deployment of a single DSVM"
author = "Graham Williams and Le Zhang"
output: rmarkdown::html_vignette
vignette: >
 %\VignetteIndexEntry{Vignette Title}
 %\VignetteEngine{knitr::rmarkdown}
 \usepackage[utf8]{inputenc}
---

# Use Case

In this tutorial, a Linux is deployed and sample code to deploy a
Windows Data Science Virtual Machine (DSVM) is provided.  Code is
included but not run to then delete the resource group if the resource
group was created within this vignette. Once deleted consumption
(cost) will cease.

This script is best run interactively to review its operation and to
ensure that the interaction with Azure completes.

As a standalone script we can run this script to setup a new resource
group and single Linux DSVM.

# Preparation

We assume the user already has an Azure subscription and we have
obtained the credentials required. See the
[Introduction](https://github.com/Azure/AzureDSVM/blob/master/vignettes/00Introduction.Rmd)
for details. We will then ensure a resource group exists and within
that resource group deploy a Linux DSVM. A secure shell (ssh) public
key matching the current user's private key is used to access the
server in this script although a username and password is also an
option. We deploy a Linux DSVM and include sample code to deploy a
Windows DSVM.

# Setup

To get started we need to load our Azure credentials as well as the
user's ssh public key. Public keys on Linux are typically created on
the users desktop/laptop machine and will be found within
~/.ssh/id_rsa.pub. It will be convenient to create a credentials file
to contain this information. The content's of the credentials file
will be something like the foloowing and we assume the user creates
such a file in the current working directory, naming the file
<USER>_credentials.R, replace <USER> with the user's username.

```{r credentials, eval=FALSE}
# Credentials come from app creation in Active Directory within Azure.
#
# See the following for details of app creation.
#
# https://github.com/Azure/AzureDSVM/blob/master/vignettes/00Introduction.Rmd
 
TID <- "72f9....db47"          # Tenant ID
CID <- "9c52....074a"          # Client ID
KEY <- "9Efb....4nwV....ASa8=" # User key

PUBKEY   <- readLines("~/.ssh/id_rsa.pub") # For Linux DSVM
PASSWORD <- "Public%4aR3@kn"               # For Windows DSVM

```

We can simply source that file in R.

```{r setup}
# Load the required subscription resources: TID, CID, and KEY.
# Also includes the ssh PUBKEY for the user.

USER <- Sys.info()[['user']]

source(paste0(USER, "_credentials.R"))
```

If the required pacakges are not yet installed the following will do
so. You may need to install them into your own local library rather
than the system library if you are not a system user.

```{r, eval=FALSE}
# Install the packages if required.

devtools::install_github("Microsoft/AzureSMR")
devtools::install_github("Azure/AzureDSVM", auth_token=GIT_TOKEN)
```

We can then load the required pacakges from the libraries.

```{r packages}
# Load the required packages.

library(AzureSMR)    # Support for managing Azure resources.
library(AzureDSVM)   # Further support for the Data Scientist.
library(magrittr)    
library(dplyr)
```

```{r tuning}
# Parameters for this script: the name for the new resource group and
# its location across the Azure cloud. The resource name is used to
# name the resource group that we will create transiently for the
# purposes of this script.

# Create a random resource group to reduce likelihood of conflict with
# other users.

runif(4, 1, 26) %>%
  round() %>%
  letters[.] %>%
  paste(collapse="") %T>%
  {sprintf("Base name:\t\t%s", .) %>% cat("\n")} ->
BASE

BASE %>%
  paste0("my_dsvm_", .,"_rg_sea") %T>%
  {sprintf("Resource group:\t\t%s", .) %>% cat("\n")} ->
RG

# Choose a data centre location.

"southeastasia"  %T>%
  {sprintf("Data centre location:\t%s", .) %>% cat("\n")} ->
LOC

# Include the random BASE in the hostname to reducely likelihood of
# conflict.

BASE %>%
  paste0("my", .) %T>%
  {sprintf("Hostname:\t\t%s", .) %>% cat("\n")} ->
HOST

cat("\n")
```

```{r connect}
# Connect to the Azure subscription and use this as the context for
# our activities.

context <- createAzureContext(tenantID=TID, clientID=CID, authKey=KEY)

# Check if the resource group already exists. Take note this script
# will not remove the resource group if it pre-existed.

rg_pre_exists <- existsRG(context, RG, LOC)

# Check that it now exists.

cat("Resource group", RG, "at", LOC,
    ifelse(!existsRG(context, RG, LOC), "does not exist.\n", "exists.\n"), "\n")
```

# Create a Resource Group

Create the resource group within which all resources we create will be
grouped.

```{r create resource group}
if (! rg_pre_exists)
{
  # Create a new resource group into which we create the VMs and
  # related resources. Resource group name is RG. 
  
  # Note that to create a new resource group one needs to add access
  # control of Active Directory application at subscription level.

  azureCreateResourceGroup(context, RG, LOC) %>% cat("\n\n")

}

# Check that it now exists.

cat("Resource group", RG, "at", LOC,
    ifelse(!existsRG(context, RG, LOC), "does not exist.\n", "exists.\n"), "\n")
```

# Deploy a Linux Data Science Virtual Machine

Create the actual Linux DSVM with public-key based authentication
method. Name, username, and size can also be configured.

Note one can check available VM sizes within the region by using
`getVMSizes()`. Different sizes will cost differently, and the
detailed information can be checked on
[Azure website](https://docs.microsoft.com/en-us/azure/virtual-machines/virtual-machines-windows-sizes). The
default VM size for deployment is chosen for by enhanced computation
performance. See the documentation for deployDSVM() for the actual
default.

```{r}
# List the available VM sizes. May differ with location of the data centre.

getVMSizes(context, "southeastasia") %>%
  set_names(c("Size", "Cores", "Disk GB", "RAM GB", "Max Disks"))

# The default size.

formals(deployDSVM)$size

# The default operating system is Ubuntu.

formals(deployDSVM)$os
```

The following code deploys a Linux DSVM, and it will take approximately 4 minutes.

```{r deploy}
# Create the required Linux DSVM - generally 4 minutes.

ldsvm <- deployDSVM(context, 
                    resource.group=RG,
                    location=LOC,
                    hostname=HOST,
                    username=USER,
                    pubkey=PUBKEY)
ldsvm

operateDSVM(context, RG, HOST, operation="Check")

azureListVM(context, RG)
```

Prove that the deployed DSVM exists.

```{r prove exists}

# Send a simple system() command across to the new server to test its
# existence. Expect a single line with an indication of how long the
# server has been up and running.

cmd <- paste("ssh -q",
             "-o StrictHostKeyChecking=no",
             "-o UserKnownHostsFile=/dev/null",
             ldsvm, "uptime")
cmd
system(cmd, intern=TRUE)

```

# Some Standard Setup

TODO: Install wajig, lsb. Then update, distupgrade. Then lsb-release.

# Deploy a Windows Data Science Virtual Machine

`deployDSVM` also supports deployment of Windows DSVM, which can be
achieved by setting the argument of `os` to "Windows". The deployment
will take approximately 10 minutes. One can use Remote Desktop to
verify the success of deployment and use the virtual machine in a
remote desktop environment.

```{r, eval=FALSE}
wdsvm <- deployDSVM(context,
                    resource.group=RG,
                    location=LOC,
                    hostname="xxxx",
                    username=USER,
                    os="Windows",
                    password=PASSWORD)

wdsvm
```

# Optional Stop

It is always a good practice to stop DSVMs after using them, to avoid any unnecessary cost.

```{r, eval=FALSE}
operateDSVM(context, RG, HOST, operation="Stop")
```

# Optional Cleanup

```{r optionally_delete_resource_group}
# Delete the resource group now that we have proved existence. There
# is probably no need to wait. Only delete if it did not pre-exist
# this script. Deletion seems to take 10 minutes or more.

if (! rg_pre_exists)
  azureDeleteResourceGroup(context, RG)
```

Once deleted we are consuming no more.
