---
title: "Using Azure Data Science Virtual Machine: Extension of DSVM"
author: "Graham Williams and Le Zhang"
date: "July 10, 2017"
output: html_document
vignette: >
 %\VignetteIndexEntry{Vignette Title}
 %\VignetteEngine{knitr::rmarkdown}
 \usepackage[utf8]{inputenc}
---

# Use case

Azure supports extensions to a VM. It is helpful in many data science application
scenarios with DSVM. For example, 

* An individual user can upgrade, downgrade, or uninstall
the pre-installed software, or install new as needed. This can be achieved after a deployment of a DSVM. 
* A custom script for system set-ups may be easily re-used across a set of DSVMs. 

More details about Azure VM Extensions
can be found [here](https://docs.microsoft.com/en-us/azure/virtual-machines/windows/extensions-features).

# Setup

```{r setup}
# Load the required subscription resources: TID, CID, and KEY.
# Also includes the ssh PUBKEY for the user.

USER <- Sys.info()[['user']]

source(paste0(USER, "_credentials.R"))
```

```{r packages}
# Load the required packages.

library(AzureSMR)    # Support for managing Azure resources.
library(AzureDSVM)    # Further support for the Data Scientist.
library(magrittr)    
library(dplyr)
```

```{r tuning}
# Parameters for this script: the name for the new resource group and
# its location across the Azure cloud. The resource name is used to
# name the resource group that we will create transiently for the
# purposes of this script.

# Create a random resource group to reduce likelihood of conflict with
# other users.

BASE <- 
  runif(4, 1, 26) %>%
  round() %>%
  letters[.] %>%
  paste(collapse="") %T>%
  {sprintf("Base name:\t\t%s", .) %>% cat("\n")}

RG <-
  paste0("my_dsvm_", BASE,"_rg_sea") %T>%
  {sprintf("Resource group:\t\t%s", .) %>% cat("\n")}

# Choose a data centre location.

LOC <-
  "southeastasia"  %T>%
  {sprintf("Data centre location:\t%s", .) %>% cat("\n")}

# Include the random BASE in the hostname to reducely likelihood of
# conflict.

HOST <-
  paste0("my", BASE) %T>%
  {sprintf("Hostname:\t\t%s", .) %>% cat("\n")}

cat("\n")
```

# Deployment of a DSVM

Deployment is the same as that in the previous sub-section. Here in the demo, a
Linux DSVM with public key type authenticaiton is deployed.

```{r connect}
context <- createAzureContext(tenantID=TID, clientID=CID, authKey=KEY)

rg_pre_exists <- existsRG(context, RG, LOC)

cat("Resource group", RG, "at", LOC,
    ifelse(!existsRG(context, RG, LOC), "does not exist.\n", "exists.\n"), "\n")
```

```{r create resource group}
if (! rg_pre_exists) {
  azureCreateResourceGroup(context, RG, LOC) %>% cat("\n\n")
}
```

```{r}
getVMSizes(context, "southeastasia") %>%
  set_names(c("Size", "Cores", "DiskGB", "RAM GB", "Disks"))

formals(deployDSVM)$size

formals(deployDSVM)$os
```

```{r deploy}
ldsvm <- deployDSVM(context, 
                    resource.group = RG,
                    location       = LOC,
                    hostname       = HOST,
                    username       = USER,
                    authen         = "Key",
                    pubkey         = PUBKEY)
ldsvm

operateDSVM(context, RG, HOST, operation="Check")

azureListVM(context, RG)
```

# Install an extension.

To install an extension to a deployed DSVM with custom script, a URL of the custom
script here it is hosted and a system command to execute that script are needed
as inputs. 

In our example, [a script](https://github.com/Azure/DataScienceVM/blob/master/Extensions/General/create-user-and-updatecntk.sh) available on Azure DSVM github repository is used. The script is to create user and update CNTK on the DSVM. The
command to execute script is "sudo sh <script.R>". 

```{r}
URL <- "https://raw.githubusercontent.com/Azure/DataScienceVM/master/Extensions/General/create-user-and-updatecntk.sh"
CMD <- "sudo sh create-user-and-updatecntk.sh"

ldsvm_ext <- addExtensionDSVM(context, 
                              resource.group = RG, 
                              location = LOC, 
                              hostname = HOST,
                              os = "Ubuntu", 
                              fileurl = URL, 
                              command = CMD)
```

A successful installation will return a TRUE. If it fails, possible reason can be
retrieved from Azure portal:

1. Go to Azure portal.
2. Click on the deployed DSVM. There is "Extensiosn" found on the left pane. 
3. Click "Extensions", and the extension installed just now can be found inside. 
4. If it failed, detailed information can be found there for troubleshooting.

Once compute resources are no longer needed, stop or delete it by 

```{r, eval=FALSE}
operateDSVM(context, RG, HOST, operation="Stop")
```

```{r optionally_delete_resource_group}
if (! rg_pre_exists)
  azureDeleteResourceGroup(context, RG)
```