---
title = "Using Azure Data Science Resources: Compute on Linux DSVM Quick Start"
author= "Graham Williams"
---

# TODO

This is currently just a copy of LinuxDSVM. Intend to submit a script
file build.R across to the server to build randomForest and then
rxDForest on weatherAUS

# Use Case

Our use case here is for a Data Scientist creating their R programs to
analyse a dataset on their local compute platform (e.g., a laptop with
6GB RAM running Ubuntu with R installed). Development is done with a
dataset size (a random sample of the full dataset perhaps) that will
not tax the available memory and will return results quickly. When the
experimental setup is complete the script can be sent across to an
considerably more capable compute engine on Azure.

A Linux Data Science Virtual Machine is deployed, the analysis
completed, results collected, and the compute resources deleted. Azure
consumption costs are minimised.

This specific demonstration simply creates a Linux Data Science
Virtual Machine within a resource group, demonstrates it exists, and
then deletes the resource group.

This script is best run interactively to review its operation and
ensure interaction with Azure completes.

# Setup

We assume there is already a subscription and we have obtained the
credentials required. See
[AzureSMR's Authentication Guide](https://github.com/Microsoft/AzureSMR/blob/master/vignettes/Authentication.Rmd)
for details.  We will then ensure a resource group exists and within
that resource group create a Linux DSVM. A public ssh key is used to
access the server.

To get started we need to load the obtained credentials as well as the
user's ssh public key. Public keys on Linux are typically created on
the users desktop/laptop machine and will be found within
~/.ssh/id_rsa.pub. The content's of the user's credentials file will
be something like:

```{r credentials, eval=FALSE}
# Credentials come from app creation in Active Directory within Azure.
 
TID <- "72f9....db47"          # Tenant ID
CID <- "9c52....074a"          # Client ID
KEY <- "9Efb....4nwV....ASa8=" # User key

PUBKEY <- readLines("~/.ssh/id_rsa.pub")
```

```{r setup}
# Load the required subscription resources: TID, CID, and KEY.
# Also includes the ssh PUBKEY for the user.

USER <- Sys.getenv("USER")

source(paste0(USER, "_credentials.R"))

# Install the packages if required.

devtools::install_github("Microsoft/AzureSMR")
devtools::install_github("Azure/AzureDSR", auth_token=GIT_TOKEN)
```

```{r packages}
# Load the required packages.

library(AzureSMR)    # Support for managing Azure resources.
library(AzureDSR)    # Further support for the Data Scientist.
library(magrittr)    
library(dplyr)
library(rattle)      # Use weatherAUS as a "large" dataset.
```

```{r tuning}
# Parameters for this script: the name for the new resource group and
# its location across the Azure cloud. The resource name is used to
# name the resource group that we will create transiently for the
# purposes of this script.

RG  <- "my_dsvm_rg_sea"  # Create if not already exist then kill.
LOC <- "southeastasia"   # Where the resource group (resources) will be hosted.
```

```{r connect}
# Connect to the Azure subscription and use this as the context for
# our activities.

context <- createAzureContext(tenantID=TID, clientID=CID, authKey=KEY)

# Check if the resource group already exists. Take note this script
# will not remove the resource group if it pre-existed.

context %>%
  azureListRG() %>%
  filter(name == RG) %>%
  select(name, location) %T>%
  print() %>%
  nrow() %>%
  equals(0) %>%
  not() %T>%
  print() ->
rg_pre_exists
```
# Creation

Create the resource group within which all resources we create will be
grouped.

```{r create resource group}
if (! rg_pre_exists)
{
  # Create a new resource group into which we create the VMs and
  # related resources. Resource group name is RG. TODO Seems that not
  # everyone can create new resource groups. Need to determine what
  # controls this. Perhaps it is a management issue and general users
  # cannot have access to manage resource groups. Instead, how do we
  # identify all the resources we create when deploying the virtual
  # machine? Do we get a list of resources in the resource group in
  # advance and compare it afterwards and simply remove all new
  # resources, assuming no one else created any in the same resource
  # group in the meantime. For now assume if prexists we simply do not
  # remove the resource group.

  azureCreateResourceGroup(context, RG, LOC)

}
```
Create the actual Linux Data Science Virtual Machine.

```{r deploy}
# Create the required Linux DSVM - generally 4 minutes.

ldsvm <- newLinuxDSVM(context, RG, LOC, "mydsvm07", USER, PUBKEY)

ldsvm
```

Prove that the server exists.

```{r prove exists}

# Send a simple system() command across to the new server to test its
# existence. Expect a single line wiht an indication of how long the
# server has been up and running.

cmd <- paste("ssh -q",
             "-o StrictHostKeyChecking=no",
             "-o UserKnownHostsFile=/dev/null",
             ldsvm, "uptime")
cmd
system(cmd)
```

# Cleanup

```{r optionally delete resource group}
# Delete the resource group now that we have proved existence. There
# is probably no need to wait. Only delete if it did not pre-exist
# this script. Deletion seems to take 10 minutes or more.

if (! rg_pre_exists) azureDeleteResourceGroup(context, RG)
```

Once deleted we are consuming no more.
